Body size patterns

### 2019-09-16 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
added model selection models

vB$speciesGG <- factor(vB$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
vB$seasonGG <- factor(vB$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

@@

vB$meanTemperatureFlowScaledBySeason <- vB$meanTemperatureScaledBySeason * vB$meanFlowScaledBySeason
```

@@

ggplot(cohortStrength %>% filter(cohort > 1999), aes(cohort, cohortStrengthScaled, group = speciesGG, shape = speciesGG, color = speciesGG)) +
  geom_point(size = 3) +
  geom_line(size = 1.2) +
  labs(x = 'Cohort',
       y = 'Cohort strength',
       shape = 'Species', color='Species') +
  theme_publication(base_size = 20) +
  scale_color_grey() +
  theme(legend.position = "none",
        #legend.position=c(0.8, 0.8),
    #legend.direction = "vertical",
    #legend.title = element_text(face = "plain"),
    strip.background = element_blank(),
    strip.text.y = element_blank()) +
  facet_grid(metaPop01 ~ .)

@@

ggplot(cohortStrengthWide %>% filter(cohort >= 2002, metaPop01 == 1), aes(cohortStrengthScaledBKT, cohortStrengthScaledBNT)) +
  geom_point(size = 4) +
  geom_smooth(method = 'lm', se = FALSE, color = 'grey30') +
  theme_publication(base_size = 24) +
  labs(x = 'Brook Trout cohort strength',
       y = 'Brown Trout cohort strength') +
  theme(legend.position = "none")

@@

@@

```{r dataIn}
dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn) %>%
  filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
```

@@

vB$riverGG <- factor(vB$riverFactor,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))

@@

@@

@@

mod2 <- bam(observedLength ~ riverFactor * factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
               s(ydayCumul, bs = 'ps'), 
               data = dataIn, 
               method = "REML")

@@

```{r remove cohort, focus on spp-specific cohort strengths}
#sppoke with keith. decided focus on cohrot is less interesting than river effects and density effects

mod1 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
             data = dataIn, method = "REML")
summary(modCS3)
#plot(modCS3, pages = 8)

@@

AIC(mod1,mod2,mod3,mod4) %>% rownames_to_column() %>% arrange(AIC)
```

@@

@@

>>>>

vB$speciesGG <- factor(vB$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
vB$seasonGG <- factor(vB$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
vB$riverGG <- factor(vB$riverFactor,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"))

@@

vB$meanTemperatureFlowScaledBySeason <- vB$meanTemperatureScaledBySeason * vB$meanFlowScaledBySeason

@@

ggplot(cohortStrength %>% filter(cohort >= 2002), aes(cohort, cohortStrengthScaled, group = speciesGG, shape = speciesGG, color = speciesGG)) +
  geom_point(size = 3) +
  geom_line(size = 1.2) +
  labs(x = 'Cohort',
       y = 'Cohort strength',
       shape = 'Species', color='Species') +
  theme_publication(base_size = 20) +
  scale_color_grey() +
  theme(legend.position = "none",
        #legend.position=c(0.8, 0.8),
    #legend.direction = "vertical",
    #legend.title = element_text(face = "plain"),
    strip.background = element_blank(),
    strip.text.y = element_blank()) +
  facet_grid(metaPop01 ~ .)

@@

ggplot(cohortStrengthWide %>% filter(cohort >= 2002, metaPop01 == 1), aes(cohortStrengthScaledBKT, cohortStrengthScaledBNT)) +
  geom_point(size = 4) +
  geom_smooth(method = 'lm', se = FALSE, color = 'grey30') +
  theme_publication(base_size = 24) +
  labs(x = 'Brook Trout cohort strength',
       y = 'Brown Trout cohort strength') +
  xlim(c(-1.4,1.9)) +
  theme(legend.position = "none")

@@

# bkt, metapop comparison
cohortStrengthWideBKT <- cohortStrengthWide %>%
  select(-cohortStrengthScaledATS,-cohortStrengthScaledBNT) %>%
  spread( metaPop01, cohortStrengthScaledBKT) %>%
  rename(metaPop0 = `0`, metaPop1 = `1`)

ggplot(cohortStrengthWideBKT %>% filter(cohort >= 2002), aes(metaPop1, metaPop0)) +
  geom_point(size = 4) +
  geom_smooth(method = 'lm', se = FALSE, color = 'grey30') +
  theme_publication(base_size = 24) +
  labs(x = 'Cohort strength connected metapopulation',
       y = 'Cohort strength isolated population') +
  xlim(c(-1.4,1.9)) +
  theme(legend.position = "none")

ggsave('./figures/cohortStrengthBKT.png', width = 11, height = 8, units='in')

CS_metaPopBKT <- lm(metaPop0 ~ metaPop1, data = cohortStrengthWideBKT %>% filter(cohort >= 2002))

summary(CS_metaPopBKT)

@@

```{r dataIn}
write.csv(vB, file = './data/out/vB.csv')

dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn) %>%
  filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))
```

@@

@@

## Count Summaries
#number of observations
vB %>% 
  filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)) %>%
  group_by( speciesGG) %>%
  summarize(n = n())

vB %>% 
  filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)) %>%
  group_by(riverGG, speciesGG, year, season) %>%
  summarize(n = n()) %>%
  as.data.frame() %>%
  filter(season == 4)

#number of individuals
tags <- vB %>% 
  filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats'), !is.na(observedLength)) %>%
  group_by(tag,species,metaPop01) %>%
  summarize(n = n())

tags %>% group_by(species) %>% summarise(n = n())
tags %>% group_by(species, metaPop01) %>% summarise(n = n())
```

@@

# across years
ggplot(vB %>% filter(cohort %in% 2000:2013,  species == 'bkt'), aes(observedLength)) +
 # geom_density(aes(color = factor(cohort)), size = 1.35) +
    geom_freqpoly(aes(color = factor(cohort)), binwidth = 8, size = 1.35) +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color = 'Cohort') +
  scale_x_continuous("Fork Length (mm)") +
  scale_y_continuous("Count") +
  #scale_color_manual(values = c("lightblue", "lightgreen")) +
  #scale_colour_brewer(palette = "Paired")+
  facet_grid(riverGG ~ ageInSamples, scales = 'free')
```

```{r length ageInSamples = 1}

ggplot(vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1), aes(yday,observedLength, color = factor(cohort))) + geom_point() + facet_grid(cohort~river)

ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1), aes(observedLength)) +
  geom_freqpoly(aes(group = cohort, color = cohort), binwidth = 7, size = 1.35) + 
  theme_publication(base_size = 20) +
  theme(
        #strip.text.x = element_text(size = 8), 
        #strip.background = element_blank(),
        #legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical"
        ) + 
  labs(color = 'Cohort') +
#  scale_color_grey(start = 0.15, end = 0.75) +
  scale_colour_gradient(low = "grey30", high = "grey90") +
  scale_x_continuous("Fork Length (mm)") +
  scale_y_continuous("Frequency") +
  facet_grid(speciesGG ~ riverGG)

ggsave('./figures/AIS1_freqpoly.png', width = 11, height = 8, units='in')

# length models for ageInSamples = 1 to look at variation within and among years/species/rivers
lenModBKT1 <- lm(observedLength ~  factor(cohort) * factor(riverGG), data = vB %>% filter(cohort %in% 2000:2013, species == 'bkt', ageInSamples == 1))
lenModBKT2 <- lm(observedLength ~  factor(cohort) + riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bkt', ageInSamples == 1))

AIC(lenModBKT1,lenModBKT2) %>% rownames_to_column() %>% arrange(AIC)
summary(lenModBKT1)

anovaBKT <- anova(lenModBKT1)
lenModBKT1ss <- anovaBKT$"Sum Sq"
print(cbind(anovaBKT,PctExp=lenModBKT1ss/sum(lenModBKT1ss)*100))

###########################
meanCVLen <- 
vB %>% filter(cohort %in% 2000:2013, ageInSamples == 1) %>%
  group_by(species, riverGG, cohort) %>%
  summarize(meanL = mean(observedLength),
            sdL = sd(observedLength))

ggplot(meanCVLen, aes(meanL, sdL, color = factor(species))) +
  geom_point() +
  facet_wrap(~riverGG)

#no visual diff among rivers in relationship btw mean and cv in len
ggplot(meanCVLen, aes(meanL, sdL, color = riverGG)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "lm")

ggplot(meanCVLen, aes(meanL, sdL, color = factor(cohort))) +
  geom_point() +
  facet_wrap(~riverGG)

#BNT
lenModBNT1 <- lm(observedLength ~  factor(cohort) * factor(riverGG), data = vB %>% filter(cohort %in% 2000:2013, species == 'bnt', ageInSamples == 1))
lenModBNT2 <- lm(observedLength ~  factor(cohort) + riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT3 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT4 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday * riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT5 <- lm(observedLength ~ 0 + factor(cohort) * riverGG * yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))

AIC(lenModBNT1,lenModBNT2) %>% rownames_to_column() %>% arrange(AIC)
summary(lenModBNT1)
anovaBNT <- anova(lenModBNT1)
lenModBNT1ss <- anovaBNT$"Sum Sq"
print(cbind(anovaBNT,PctExp=lenModBNT1ss/sum(lenModBNT1ss)*100))

# both species together - use this for paper

lenMod1 <- lm(observedLength ~  factor(cohort) * factor(riverGG) * speciesGG, data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1))
lenMod2 <- lm(observedLength ~  (factor(cohort) + factor(riverGG) + speciesGG)^2, data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1))
lenMod3 <- lm(observedLength ~  factor(cohort) + factor(riverGG) + speciesGG, data = vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt','bnt'), ageInSamples == 1))

AIC(lenMod1,lenMod2,lenMod3) %>% rownames_to_column() %>% arrange(AIC)

summary(lenMod1)
anova1 <- anova(lenMod1)
lenMod1ss <- anova1$"Sum Sq"
print(cbind(anova1,PctExp=lenMod1ss/sum(lenMod1ss)*100))

calc.relimp(lenMod3)
#                         lmg
# factor(cohort)  0.193090771
# factor(riverGG) 0.045329273
# speciesGG       0.000531708
```

```{r length at ageInSamples1 vs cohort strength}

# NO relationship between cohortStrengthScaled and meanLen for ageInSamples = 1

ais1 <- vB %>% filter(cohort %in% 2000:2013, ageInSamples == 1) %>%
  group_by(speciesGG, metaPop01, cohort) %>%
  summarise(meanLen = mean(observedLength),
            meanYday = mean(yday))

cs_ais1 <- left_join(cohortStrength,ais1)

# does mean size increase with day of year? No, but there are some early samples with yday < 255 that have small meanLen
ggplot(cs_ais1, aes(meanYday,meanLen, color = species)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  facet_grid(~ metaPop01)
summary(lm(meanLen ~ 0 + meanYday * species * factor(metaPop01), data = cs_ais1))

ggplot(cs_ais1 %>% filter(meanYday > 0, species %in% c('bkt','bnt')), aes(cohortStrengthScaled, meanLen)) +
  geom_point(aes(size = meanYday)) +
  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color = 'Cohort') +
  scale_y_continuous("Mean Fork Length (mm)") +
  scale_x_continuous("Cohort Strength") +
  facet_grid(speciesGG ~ metaPop01)

ggsave('./figures/cohortStrengh_v_meanLenAIS1.png', width = 11, height = 8, units='in')

summary(lm(meanLen ~  cohortStrengthScaled * factor(metaPop01), data = cs_ais1 %>% filter(species == 'bkt')))

summary(lm(meanLen ~  cohortStrengthScaled, data = cs_ais1 %>% filter(species == 'bnt', metaPop01 == 1)))
```

@@

mod2 <- bam(observedLength ~ riverFactor * factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

@@

```{r remove cohort, focus on spp-specific cohort strengths}
#sppoke with keith. decided focus on cohort is less interesting than river effects and density effects

mod1 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
               s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
             data = dataIn, method = "REML")
#summary(modCS3)
#plot(modCS3, pages = 8)

@@

mod5 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) ,
               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
             data = dataIn, method = "REML")

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) ,
          #     s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
             data = dataIn, method = "REML")

mod7 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') ,
            #   s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) ,
          #     s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
             data = dataIn, method = "REML")

mod8 <- bam(observedLength ~ riverFactor + factor(salmon01) ,              
             #  s(ydayCumul, bs = 'cr') 
            #   s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) ,
          #     s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               #s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
         #      s(cohort, bs = 'cr', k = kIn, by = riverFactor),
              # s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = riverFactor) +
              # s(cohortStrengthScaledBNT2p5, bs = 'cr', k = kIn, by = riverFactor), 
  #             s(cohortStrengthScaledATS2p5, bs = 'cr', k = kIn),
             data = dataIn, method = "REML")

AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7,mod8) %>% rownames_to_column() %>% arrange(AIC)
```

@@

# modBKT2 <- bam(observedLength ~ riverFactor + factor(salmon01) +              
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) +
#                s(cohortStrengthScaledBKT,cohortStrengthScaledBNT2p5, k = kIn, by = riverFactor) ,
#             
#              data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")
# 
# AIC(modBKT,modBKT2)

@@

# across bkt cohort strengths
ggplot(predGrid2Both %>% 
         filter(salmon01 == 0, cohort == 2006, cohortStrengthScaledBNT2p5 == 0),
       aes(ydayCumul, pred, group = cohortStrengthScaledBKT, color = cohortStrengthScaledBKT)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Brook Trout cohort strength') +
  #scale_color_viridis_c() +
  theme_publication(base_size = 20) +
  theme(legend.position = c(0.7, 0.37), 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 16)) +
  facet_grid(riverGG ~ speciesGG)

ggsave('./figures/ydayCumulAcrossCSBKTByRiver.png', width = 8, height = 4.75, units='in')

# across bnt cohort strengths
ggplot(predGrid2Both %>% 
         filter(salmon01 == 0, cohort == 2006, cohortStrengthScaledBKT == 0,
                riverGG %in% c('West Brook', 'Open Large')),
       aes(ydayCumul, pred, group = cohortStrengthScaledBNT2p5, color = cohortStrengthScaledBNT2p5)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Brown Trout cohort strength') +
  #scale_color_viridis_c() +
  theme_publication(base_size = 20) +
  theme(legend.position = c(0.7, 0.37), 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 16)) +
  facet_grid(riverGG ~ speciesGG)

ggsave('./figures/ydayCumulAcrossCSBNTByRiver.png', width = 8, height = 4.75, units='in')
